### 2009-04-23 doc/storage_system/arc-storage-documentation.tex
implementing requested changes in internal review comments

The file replicas can be in the following states: the replica can be valid and alive or just in the process of creation or it may be corrupt or a whole storage node may be offline. This state is always stored as metadata of the given replica. For each file there is a checksum calculated, and this checksum is used to detect if a replica gets corrupted. If a storage node (more precisely: the Shepherd service on the storage node) detects that a file is invalid, it reports this so the metadata will be in sync with the real state. And the storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat also triggers the modification of metadata.

@@

\begin{enumerate}
    \item The client tool connects to a Bartender and sends a `getFile' request with the LN of the file.
    \item The Bartender contacts a Librarian to traverse the Logical Name and to get the metadata of our file.
    \item The Librarian does the step-by-step traversing of the name space and gets all the metadata from the A-Hash and returns the metadata of our file to the Bartender. This metadata contains the location of the file's replicas and the access policies of this file.
    \item The Bartender checks, based on the access policies and the user's identity, if he is allowed to get the file, and if allowed, it chooses a replica location (currently randomly). A location consists of the URL of a Shepherd service, and the ID of the replica within that Shepherd (which is called a `referenceID'). The Bartender sends a `get' request to the chosen Shepherd.
    \item The Shepherd prepares the file transfer by asking the storage element service to create a new one-time TURL for this replica. The Shepherd returns the TURL to the Bartender.
    \item The Bartender returns the TURL to the client tool.
    \item The client tool use this TURL to get the file directly from the storage element service on the storage node.
\end{enumerate}

@@

The user has a file on his local disk and wants to upload the file to a collection called \verb#/ourvo/common/docs#. (See Figure~\ref{fig:uploading}.)
\begin{enumerate}
    \item The client tool contact a Bartender to upload the file, providing the file's metadata, including the the Logical Name (LN), which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
    \item The Bartender asks a Librarian to traverse this LN.
    \item The Librarian traverses the Logical Name, and stopping at the \verb#/ourvo/common/docs# part of the LN, which means that the name is available and the parent collection exists. If everything is OK, the metadata of the parent collection is returned to the Bartender. 
    \item The Bartender checks the access policies to decide if the user has permissions to put anything into this collection. Then it asks the Librarian to create a new file entry. 
    \item The Librarian creates the entry and returns its GUID.
    \item Next, the Bartender adds the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf#. However this LN points to a file which has currently no replica at all. If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, they would get an error message.
    \item The Bartender asks the Librarian about Shepherd services (which are sitting on storage nodes).
    \item The Librarian returns a list of Shepherd services.
    \item The Bartender randomly chooses a Shepherd service and sends it a `put' request to initiate the file upload. 
    \item The Shepherd communicates with the storage element service on the same node to create a new transfer URL (TURL). Then it creates a `referenceID' for this file and reports to the Librarian that there is a new replica in the \verb#creating# state. The Librarian gets the message from the Shepherd and adds the new replica to the new file. Now the file has one replica, which is not uploaded yet into the system. If someone tries to download this file now, they still get an error message.
    \item The Shepherd returns the the TURL to the Bartender.
    \item The Bartender returns the TURL to the client tool.
    \item Then client tool can upload the file to this TURL.
    \item The Shepherd detects that the file has been uploaded. In order to do this the storage element service does post processing (e.g. remove a hardlink, do a SOAP call, etc.) after it got the file. The Shepherd checks the checksum of the file, and if it is OK, then it reports to the Librarian, that this replica is now \verb#alive#. The Librarian alters the state of this location, and now the file has one valid replica. 
\end{enumerate}

@@

The HED has a security framework which the Bartenders use the make access policy decisions. The details of this is described in Section \ref{sec:security_implementation}.

@@

A file in Chelonia has a logical entry in the namespace, and one or more physical replicas on the storage nodes. We store the locations of the replicas in the \emph{locations} section. A location contains the URL of the Shepherd service, the local ID of the file within the storage node and the state of the replica. This state is `\textbf{alive}' if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy, `\textbf{invalid}' if the replica has wrong checksum, or the Shepherd claims it has no such file, `\textbf{offline}' if the Shepherd is not reachable, but may have a valid replica, `\textbf{creating}' if the replica is in the state of uploading, `\textbf{thirdwheel}' if the replica is marked for deletion because of too many replicas, or `\textbf{stalled}' if the the replica was in the state if uploading for a long time but hasn't arrived. The \emph{states} section contains the size of the file, the number of required replicas, and a checksum of the file, which is created by some kind of checksumming algorithm (currently only md5 is supported).

@@

A \emph{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-hierarchy. Each entry has a unique name within a collection and a GUID which points to a corresponding file or collection, so a collection is basically a list of name-GUID pairs, which list is stored in the \emph{entries} section (see below). The collections can be in a closed state, which means that its contents should not be changed. If you close a collection then it cannot be opened again. However it cannot be guaranteed that the contents of a closed collection remains the same (e.g. if we include in our collection a file which we don't own, then the owner of the file can still remove it), but if a closed collection is changed, its state will be broken, and this state can never be changed again, which means you will always know if a closed collection is not intact anymore. This state is stored in the \emph{states} section, and its possible values are `\textbf{no}' if the collection is not closed, `\textbf{yes}' if the collection is closed and `\textbf{broken}' if the collection was closed, but something happened, and its contents have been changed.

@@

\begin{description}
    \item[AHashClass] tells the A-Hash service the name of the business-logic A-Hash class. Currently there are two implementation, the \verb!CentralAHash! is a centralized version, and the \verb!ReplicatedAHash! is a replicated version.
    \item[StoreClass] specifies the name of the store class, used to store the data on disk. Currently there are several store classes each has the same interface but uses different mechanisms (\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore! class is suitable for use in a replicated deployment. All the others are only for a centralized deployement, so this variable is only configurable for the centralized deployement. Based on our tests, currently the \verb!CachedPickleStore! has the best performance.
    \item[StoreCfg] contains parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash can save its files. This variable is only configurable for the centralized deployment.
\end{description}

@@

If a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), the Shepherd service creates a hardlink into the transfer directory (\verb!./transfer/abc!) and sets this file read-only. If the Hopi service is configured to serve the HTTP path \verb!/prb! and it is serving files from the directory \verb!./transfer! then after the hardlink is created, the transfer URL for this file reads \verb!http://localhost:60000/prb/abc!. Now this URL is passed to the client. Then the client \verb!GET!'s this URL and gets the file. The Hopi service removes (unlinks) this file immediately after the \verb!GET! request arrived, which makes \verb!http://localhost:60000/prb/abc! invalid (so this is a one-time URL). However, because of the hardlink the file is still in the store directory, it is just removed from the transfer directory. Now if some other user wants this file, the Shepherd creates an other hardlink, e.g., \verb!./transfer/qwe! with the URL \verb!http://localhost:60000/prb/qwe!.

@@

The Bartender utilizes gateway modules to access file listings and transfer URL's from third-party storage systems. It works as an optional backend component of the Bartender service. The \emph{Gateway} module is based on four basic methods: \emph{get}, \emph{put}, \emph{list} and \emph{remove}. This module should connect to the third-party storages on behalf of the user, so the user has to delegate his or her credentials to the Bartender using proxy delegation. In order to access the files and directories on the external storage system, limited user credentials are passed to the Bartender service. Users can also remove their credentials from the service side. 
% subsection implementation (end)

@@

The credentials of the user and the URL of the Bartender must be configured either with an XML file located at \verb!~/.arc/client.xml! or with environment variables.

>>>>

The file replicas can be in the following states: the replica can be valid and alive or just in the process of creation or it may be corrupt or a whole storage node may be offline. This state is always stored as metadata of the given replica. For each file there is a checksum calculated, and this checksum is used to detect if a replica gets corrupted. If a storage node (more precisely: the Shepherd service on the storage node) detects that a file is invalid, it reports this so the metadata will be in sync with the real state. The storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat also triggers the modification of metadata.

@@

\begin{enumerate}
    \item The client tool connects to a Bartender and sends a `getFile' request with the LN of the file.
    \item The Bartender contacts a Librarian to traverse the Logical Name and to get the metadata of our file.
    \item The Librarian does the step-by-step traversing of the name space and gets all the metadata from the A-Hash and returns the metadata of our file to the Bartender. This metadata contains the location of the file's replicas and the access policies of this file.
    \item The Bartender checks, based on the access policies and the user's identity, if he is allowed to get the file, and if allowed, it chooses a replica location (currently randomly). A location consists of the URL of a Shepherd service, and the ID of the replica within that Shepherd (which is called a `referenceID'). The Bartender sends a `get' request to the chosen Shepherd.
    \item The Shepherd prepares the file transfer by asking the storage element service to create a new one-time TURL for this replica. The Shepherd returns the TURL to the Bartender.
    \item The Bartender returns the TURL to the client tool.
    \item The client tool use this TURL to get the file directly from the storage element service (SES) on the storage node.
\end{enumerate}

@@

The user has a file on his local disk and wants to upload the file to a collection called \verb#/ourvo/common/docs#. (See Figure~\ref{fig:uploading}.)
\begin{enumerate}
    \item The client tool contact a Bartender to upload the file, providing the file's metadata, including the the Logical Name (LN), which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
    \item The Bartender asks a Librarian to traverse this LN.
    \item The Librarian traverses the Logical Name, and stopping at the \verb#/ourvo/common/docs# part of the LN, which means that the name is available and the parent collection exists. If everything is OK, the metadata of the parent collection is returned to the Bartender. 
    \item The Bartender checks the access policies to decide if the user has permissions to put anything into this collection. Then it asks the Librarian to create a new file entry. 
    \item The Librarian creates the entry and returns its GUID.
    \item Next, the Bartender adds the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf#. However this LN points to a file which has currently no replica at all. If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, they would get an error message.
    \item The Bartender asks the Librarian about Shepherd services (which are sitting on storage nodes).
    \item The Librarian returns a list of Shepherd services.
    \item The Bartender randomly chooses a Shepherd service and sends it a `put' request to initiate the file upload. 
    \item The Shepherd communicates with the storage element service (SES) on the same node to create a new transfer URL (TURL). Then it creates a `referenceID' for this file and reports to the Librarian that there is a new replica in the \verb#creating# state. The Librarian gets the message from the Shepherd and adds the new replica to the new file. Now the file has one replica, which is not uploaded yet into the system. If someone tries to download this file now, they still get an error message.
    \item The Shepherd returns the the TURL to the Bartender.
    \item The Bartender returns the TURL to the client tool.
    \item Then client tool can upload the file to this TURL (to the storage element service, SES)
    \item The Shepherd detects that the file has been uploaded. In order to do this the storage element service does post processing (e.g. remove a hardlink, do a SOAP call, etc.) after it got the file. The Shepherd checks the checksum of the file, and if it is OK, then it reports to the Librarian, that this replica is now \verb#alive#. The Librarian alters the state of this location, and now the file has one valid replica. 
\end{enumerate}

@@

The HED has a security framework which the Bartenders use to make access policy decisions. The details of this is described in Section \ref{sec:security_implementation}.

@@

A file in Chelonia has a logical entry in the namespace, and one or more physical replicas on the storage nodes. We store the locations of the replicas in the \emph{locations} section. A location contains the URL of the Shepherd service, the local ID of the file within the storage node and the state of the replica. This state is `\textbf{alive}' if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy, `\textbf{invalid}' if the replica has wrong checksum, or the Shepherd claims it has no such file, `\textbf{offline}' if the Shepherd is not reachable, but may have a valid replica, `\textbf{creating}' if the replica is in the state of uploading, `\textbf{thirdwheel}' if the replica is marked for deletion because of too many replicas, or `\textbf{stalled}' if the the replica was in the state of uploading for a long time but hasn't arrived. The \emph{states} section contains the size of the file, the number of required replicas, and a checksum of the file, which is created by some kind of checksumming algorithm (currently only md5 is supported).

@@

A \emph{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-like hierarchy (with possible interconnection due to hardlinks). Each entry has a unique name within a collection and a GUID which points to a corresponding file or collection, so a collection is basically a list of name-GUID pairs, which list is stored in the \emph{entries} section (see below). The collections can be in a closed state, which means that its contents should not be changed. If you close a collection then it cannot be opened again. However it cannot be guaranteed that the contents of a closed collection remains the same (e.g. if we include in our collection a file which we don't own, then the owner of the file can still remove it), but if a closed collection is changed, its state will be broken, and this state can never be changed again, which means you will always know if a closed collection is not intact anymore. This state is stored in the \emph{states} section, and its possible values are `\textbf{no}' if the collection is not closed, `\textbf{yes}' if the collection is closed and `\textbf{broken}' if the collection was closed, but something happened, and its contents have been changed.

@@

\begin{description}
    \item[AHashClass] tells the A-Hash service the name of the business-logic A-Hash class. Currently there are two implementations, the \verb!CentralAHash! is a centralized version, and the \verb!ReplicatedAHash! is a replicated version.
    \item[StoreClass] specifies the name of the store class, used to store the data on disk. Currently there are several store classes each has the same interface but uses different mechanisms (\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore! class is suitable for use in a replicated deployment. All the others are only for a centralized deployement, so this variable is only configurable for the centralized deployement. Based on our tests, currently the \verb!CachedPickleStore! has the best performance.
    \item[StoreCfg] contains parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash can save its files. This variable is only configurable for the centralized deployment.
\end{description}

@@

If a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file1!), the Shepherd service creates a hardlink into the transfer directory (\verb!./transfer/abc!) and sets this file read-only. If the Hopi service is configured to serve the HTTP path \verb!/prb! and it is serving files from the directory \verb!./transfer! then after the hardlink is created, the transfer URL for this file reads \verb!http://localhost:60000/prb/abc!. Now this URL is passed to the client. Then the client \verb!GET!'s this URL and gets the file. The Hopi service removes (unlinks) this file immediately after the \verb!GET! request arrived, which makes \verb!http://localhost:60000/prb/abc! invalid (so this is a one-time URL). However, because of the hardlink the file is still in the store directory, it is just removed from the transfer directory. Now if some other user wants this file, the Shepherd creates an other hardlink, e.g., \verb!./transfer/qwe! with the URL \verb!http://localhost:60000/prb/qwe!.

@@

The Bartender utilizes \emph{Gateway} modules to access file listings and transfer URLs from third-party storage systems. The gateway module has four basic methods: \emph{get}, \emph{put}, \emph{list} and \emph{remove}. This module connects to the third-party storages on behalf of the user, in order to do this, the Bartender needs to have proxy certificates from the user. A proxy certificate can be passed to the Bartender through the process of credential delegation. Users can also remove their proxy certificate from the service side. 
% subsection implementation (end)

@@

The credentials of the user and the URL of the Bartender must be configured either with an XML file located at \verb!~/.arc/client.xml! or with environment variables. (The environment variables overwrite the values from the XML file.)
